---
title: "Homework 10"
author: "T.J. Seislove"
date: "4/7/2026"
output: 
  html_document:
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(mosaic)
library(janitor)
library(knitr)
library(lubridate)
```

# General Social Survey

### Data Import

```{r}
data(gss_cat)
```

### Marital Levels

```{r}
gss_marital <- gss_cat %>% mutate(marital = fct_collapse(marital,
                         "Pre-Marriage" = c("Never married"),
                         "Marriage" = c("Married", "Separated"),
                         "Post-Marriage" = c("Divorced", "Widowed"),
                          "Missing" = c("No answer")
                        ))

gss_marital %>% count(marital) %>% kable()
```

^[Problem 1 Complete]

### Year and Marrital Status

```{r}
marriage_year <- gss_marital %>% count(year, marital) %>% 
  group_by(year) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()

head(marriage_year) %>% kable()
```

^[Problem 2 Complete]

```{r}
marriage_year <- marriage_year %>% mutate(
  marital = fct_reorder2(marital, year, prop)
)
```

^[Problem 3 Complete]

### Line Plot

```{r}
marriage_year %>% ggplot(mapping = aes(x = year, y = prop, color = marital)) +
  geom_line()
```

^[Problem 4 Complete]

# Los Angeles Lakers

### Dataset Import

```{r}
data("lakers")
```

### Converting Date Data

**Integer --> Date**

```{r}
lakers_date <- lakers %>% mutate(
  date_data = as_date(date),
  .keep = "used"
)

head(lakers_date) %>% kable()
```

I think the conversion failed because as_date() doesn't parse the integer into a separate year, month, and day. It looks like it interprets the integer as a number of days since some starting point. Thus it thinks the year is beyond 56,900.

^[Problem 1 Complete]

**Integer --> String --> Date**

```{r}
lakers_date <- lakers %>% mutate(date = as.character(date)) %>% 
  mutate(date_data = as_date(date), .keep = "used")

head(lakers_date) %>% kable()
```

I think this works better because as_date() does parse through the string and it finds separate year, month, and day values. It interprets the first 4 characters as the year, the next 2 as the month, and the last 2 as the day, as opposed to treating the whole string as one value like it did with the integer.

^[Problem 2 Complete]

### Days of the Week

**Find Distinct Dates**

```{r}
lakers_distinct <- distinct(lakers_date)

head(lakers_distinct) %>% kable()
```

**Extract Day of the Week**

```{r}
lakers_distinct <- lakers_distinct %>% mutate(
  wday = wday(date_data, label = TRUE)
)

head(lakers_distinct) %>% kable()
```

**Make Barplot**

```{r}
ggplot(data = lakers_distinct, aes(x = wday, fill = wday)) +
  geom_bar()
```

^[Problem 3 Complete]